---
title: "Anova analysis"
author: "Giovanna Morara"
output: html_notebook
---

```{r}
# Clear the environment
rm(list = ls())

# Load the libraries
library(here)

# Load data
data <- read.csv(here("Data", "hsb2.csv"))
```

# Describing Data

```{r}
# ses ordinal  (1=low 2=middle 3=high)  
data$ses <- factor(data$ses,
                   labels = c("Low",
                              "Middle",
                              "High"))

# Obtain boxplots.

boxplot(data$science ~ data$ses,
        ylim = c(0, 100),
        xlab = "Socioeconomic Status",
        ylab = "Science Scores",
        main = "Scores depending on SES")

```

## Obtain descriptive statistics.

```{r}
tapply(data$science, data$ses, summary)
tapply(data$science, data$ses, sd)
```


# Check the conditions necessary for valid ANOVA inference

### Groups are independent (research design)

This study is observational so individuals are classified to one SES.

### Units within groups are independent to each other (research design)

The students are randomly selected. It could still be possible that near
relatives are from one family (therefore non independent).

### Now let's check the conditions for inference (statistic)

```{r}
model <- lm(data$science ~ data$ses)
std_residual <- rstandard(model)
qqnorm(std_residual)
qqline(std_residual)
```

The sampling distribution is normally distributed, although there is 
a deviation for higher levels. However, for each group we have more
than 30 participants and the deviations is not harsh. Therefore, the 
condition is met.

```{r}
table(data$ses)
```

### Residual withint groups have the same variation

Looking at the variance of residuals within the categories. 

```{r}
plot(model$fitted.values, model$residuals,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Plot of Residuals vs. Fit")

abline(0,0)
```

There is some skewness in the in the High SES. But the spread around the 
means homogeneous for the most cases.

### Sampling distribution of means are normal distributions

We do not have a skewed distribution, we also have more than 30
participant for each group so for the Central Limit Theorem, so we have
a normal sampling distribution. 

---------------------------------------------------------------------------

# Conduct an analysis of variance (ANOVA) for omnibus inference

We will use function aov to get the analysis of variance

```{r}
aov.model <- aov(data$science ~ data$ses)
aov.model
summary(aov.model)
```
Choosing a 95% certanty level, we can reject the null hypothesis: there is
some difference between one or more of the groups.

# Anlalysis of pairs of mean.

This section is to construct confidence intervals for each pair of 
conditions to understand the differences between the groups.

```{r}
## Checking dummies variables
# model <- lm(data$science ~ data$ses)
# summary(model)
# anova(model)
```


```{r}
t.test(science ~ ses,
           data = data,
           subset = (ses == "Low" | ses == "Middle"))

t.test(science ~ ses,
           data = data,
           subset = (ses == "Middle" | ses == "High"))

t.test(science ~ ses,
           data = data,
           subset = (ses == "Low" | ses == "High"))
```

There seems to be differences for each pair of means. Specifically, science
scores for high ses students are higher then both middle and low ses. At 
the same time, middle ses students have higher science score than low ses
students. 